---
title: "Zone 2 Deer Forecast"
output: html_document
params:
  initPrms: NA
  TotPop: NA
  TotPopGraph: NA
  prev1: NA
  prev2: NA
  PrevsGraph: NA
  PrevMap: NA
  grid: NA
  PrevMapAlt: NA
  Lambda: NA
  Age: NA
  Densplot: NA
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE) 
```

## Simulation Parameters

`r params$initPrms`

## Density

```{r}
TotDens <- params$TotPop
maxyear <- max(TotDens$t)
TotDens2 <- subset(TotDens, TotDens$group == 'All' & TotDens$t %in% c(1, maxyear))

```

In year 1, the average deer density in Zone 2 was `r round(TotDens2$N[1], 2)`. By year `r maxyear`, the average deer density is predicted to be `r round(TotDens2$N[2], 2)`. 

```{r}
params$TotPopGraph
```

```{r}
density <- params$Densplot/6.158
grid0_plot <- params$grid
for(j in c(1, maxyear)){
grid0_plot$Density <- density[j,]
grid0_plot$Density[stoch_dd_s$noGo] <- NA
county_levels <- as.data.frame(grid0_plot %>%
                  group_by(County) %>%
                  summarize(Density = mean(Density, na.rm = T)))[c(1:3,6:7,9:10),]
county_levels$Density <- round(county_levels$Density)
county_levels$x <- c(561570.5,489570.5,459570.5,
                           531470.5,481570.5,520570.5,570570.5)
county_levels$y <- c(3997732, 4023732, 4031732,
                           4005732,3981732,3975732, 3977732)

print(ggplot() +
        geom_tile(data= grid0_plot, aes(x = x, y = y, fill = Density), alpha = .8) +
        #scale_fill_viridis_d(option = 'B', alpha = .8, limits = unique(levs))+
        scale_fill_gradient2(low = 'black', mid  = 'pink',midpoint = .25, high = 'firebrick',
                             na.value = 'white',
                             limits = c(0,max(params$Densplot, na.rm = T)/6.178 + .4)
                             )+
        theme_minimal() +
        geom_label(data = county_levels, aes(x = x, y=y, label = Density))+
        theme(
          strip.text = element_text(size = 10),  # Adjust facet labels
          title = element_text(size = 20),
          axis.title = element_blank(),
          legend.text = element_text(size =18),
          axis.text = element_blank(),
          panel.grid = element_blank()
        )+
        ggtitle(paste0("Deer Density per Sq Mile in Time ", j))
)
}
```

## Population Growth Rate

```{r}
params$Lambda
```


## Age Ratios

```{r}
simpleage <- params$Age
finalyear <- subset(simpleage, simpleage$t == maxyear)
```

In year `r maxyear`, the predicted age ratio of the population is predicted to be `r round(finalyear$N[1]*100, 2)`% fawns, `r round(finalyear$N[2]*100, 2)`% yearlings, and `r round(finalyear$N[3]*100, 2)`% adults. 

```{r}
ggplot(simpleage, aes(x = t, y = N, group = group, fill = group))+
  geom_col(position = position_dodge2(preserve = 'single', width = .8), col = 'black', lwd = .8)+
  theme_bw()+
  scale_fill_manual(values = (c('#30586C', '#549EC3','#6BC9E3')))+
  theme(axis.text = element_text(size = 20),
        axis.ticks = element_blank(),
        axis.title = element_text(size = 22),
        legend.text = element_text(size = 18),
        legend.title = element_text(size = 20),
        strip.text = element_text(size = 20),
        title = element_text(size = 20))

```


## CWD Prevalence 
```{r}
prevs_avg <- params$prev1
prevs_avgAds <- subset(prevs_avg, prevs_avg$group == 'Adult Prevalence')
prevs_avgAll <- subset(prevs_avg, prevs_avg$group != 'Adult Prevalence')
prevs_bypix <- params$prev2
prevs2 <- subset(prevs_bypix, prevs_bypix$t %in% c(1, maxyear))
```

In year 1, average prevalence across Zone 2 (including fawns) was `r round(prevs_avgAll$Prop[1], 2)`. Adult prevalence (only including yearlings and adults) was `r round(prevs_avgAds$Prop[1], 2)`. 

In year `r maxyear`, average prevalence is predicted to be `r round(prevs_avgAll$Prop[maxyear], 2)` for all deer and `r round(prevs_avgAds$Prop[maxyear], 2)` for adult deer. 

Prevalence in year `r maxyear` ranges from `r round(range(prevs2$Prop)[1],2) ` to `r round(range(prevs2$Prop)[2],2) ` depending on location. 


```{r}
params$PrevsGraph
```

Here prevalence refers to adult prevalence, as this number will most closely reflect test results and sample prevalence. 

```{r}
grid0_plot <- params$grid
prevmap <- params$PrevMap
for(t in c(1, maxyear)){
myyear <- subset(prevmap, prevmap$time ==  t)
grid0_plot$Prevalence <- myyear$AvgPrev
grid0_plot$Prevalence[stoch_dd_s$noGo] <- NA
county_levels <- as.data.frame(myyear %>% 
                               group_by(county) %>% 
                               summarize(Prev = mean(AvgPrev, na.rm = T)))[c(1:3,6:7,9:10),]
county_levels$Prev <- round(county_levels$Prev, 2)
county_levels$x <- c(561570.5,489570.5,459570.5,
                      531470.5,481570.5,520570.5,570570.5)
county_levels$y <- c(3997732, 4023732, 4031732,
                      4005732,3981732,3975732, 3977732)
     
print(ggplot() +
        geom_tile(data= grid0_plot, aes(x = x, y = y, fill = Prevalence), alpha = .8) +
         scale_fill_viridis_c(option = 'B', alpha = .8, direction = 1,
                              limits = c(0,max(prevmap$AvgPrev, na.rm = T) + .1))+
        theme_minimal() + 
        theme(
          strip.text = element_text(size = 10),  # Adjust facet labels
          title = element_text(size = 15),
          legend.text = element_text(size =18),
          axis.title = element_blank(),
          axis.text = element_blank(),
          panel.grid = element_blank()
        )+
        #geom_spatvector(data = counties, fill = NA, col = 'grey80', lwd = .5, lty = 2)+
        #geom_spatvector(data = zone2, fill = NA, col = 'grey80', lwd = 1)+
        geom_label(data = county_levels, aes(x = x, y=y, label = Prev))+
        ggtitle(paste0("Predicted CWD Prevalence at Time ", t))
)
}
```


```{r}
# for(t in c(maxyear)){
# myyear <- subset(params$PrevMapAlt, params$PrevMapAlt$time ==  t)
# grid0_plot$Prevalence <- myyear$prev
# grid0_plot$Prevalence[stoch_dd_s$noGo] <- NA
#      
# print(ggplot() +
#         geom_tile(data= grid0_plot, aes(x = x, y = y, fill = Prevalence), alpha = .8) +
#          scale_fill_viridis_c(option = 'A', alpha = .8, direction = 1,
#                               limits = c(0,max(params$PrevMapAlt$prev, na.rm = T) + .1))+
#         theme_minimal() + 
#         theme(
#           strip.text = element_text(size = 10),  # Adjust facet labels
#           title = element_text(size = 15),
#           legend.text = element_text(size =18),
#           axis.title = element_blank(),
#           axis.text = element_blank(),
#           panel.grid = element_blank()
#         )+
#         #geom_spatvector(data = counties, fill = NA, col = 'grey80', lwd = .5, lty = 2)+
#         #geom_spatvector(data = zone2, fill = NA, col = 'grey80', lwd = 1)+
#         #geom_label(data = county_levels, aes(x = x, y=y, label = Prev))+
#         ggtitle(paste0("Predicted CWD Prevalence at Time ", t))
# )
# }
```

## Infection Model:

The infection model takes into account both the density of infected deer in the current year (x-axis) and the density of infected deer in the previous year (panels)


```{r}
eta_d <- function(x, y){.25*(1-exp(-.974*(.75*x + .25*y)))} #x= current pos dens, y = prev
curr_inf <- seq(0, 5, by = .1)
prev_inf <- c(0, 1, 3, 5)
inf <- array(NA, c(length(curr_inf), length(prev_inf)))
for(j in 1:length(curr_inf)){
  for(k in 1:length(prev_inf)){
    inf[j,k] <- eta_d(curr_inf[j], prev_inf[k])
  }
}

gg_inf <- data.frame(inf = c(inf),
                     current = rep(curr_inf, length(prev_inf)),
                     `Previous Infected Density` = rep(prev_inf, each= length(curr_inf)))
gg_inf$`Previous Year's Infected Density` <- as.factor(gg_inf$Previous.Infected.Density)
ggplot(gg_inf, aes(x = current, y = inf, col = `Previous Year's Infected Density`))+
  geom_line()+
  facet_wrap(~`Previous Year's Infected Density`)+
  theme_bw()+
  ylab("Infection Probability")+
  xlab("Infected Deer Density (per square mile)")+
  theme(strip.text = element_text(size = 15),  # Adjust facet labels
        axis.text = element_text(size = 15),
        legend.text = element_text(size =10),
        legend.title =element_text(size = 12) )

```


Report created:  `r Sys.time()` 

